---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cpsR

<!-- badges: start -->

[![CRAN status](https://www.r-pkg.org/badges/version/cpsR)](https://CRAN.R-project.org/package=cpsR)
[![R-CMD-check](https://github.com/matt-saenz/cpsR/workflows/R-CMD-check/badge.svg)](https://github.com/matt-saenz/cpsR/actions)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/grand-total/cpsR)](https://cran.r-project.org/package=cpsR)

<!-- badges: end -->

## Overview

Load [Current Population Survey (CPS)](https://www.census.gov/programs-surveys/cps/about.html) microdata into R using the Census Bureau Data API, including [basic monthly CPS](https://www.census.gov/data/datasets/time-series/demo/cps/cps-basic.html) and [CPS ASEC](https://www.census.gov/data/datasets/time-series/demo/cps/cps-asec.html) microdata.

Note: This product uses the Census Bureau Data API but is not endorsed or certified by the Census Bureau.

## Installation

To install cpsR, run the following code:

``` r
install.packages("cpsR")
```

To install the development version of cpsR, run the following code:

``` r
# install.packages("devtools")
devtools::install_github("matt-saenz/cpsR")
```

## Census API key

In order to use cpsR functions, you must supply a [Census API key](https://api.census.gov/data/key_signup.html) in one of two ways:

1. Using the `key` argument (manually)
2. Using environment variable `CENSUS_API_KEY` (automatically)

Using environment variable (or env var, for short) `CENSUS_API_KEY` is strongly recommended for two reasons:

1. Saves you from having to copy-paste your key around
2. Allows you to avoid including your key in scripts

Number two is particularly important if you plan to share your code with others (like in the [example](#example) below) since you should keep your key secret.

You can set up env var `CENSUS_API_KEY` in two steps:

First, open your `.Renviron` file. You can do so by running:

``` r
# install.packages("usethis")
usethis::edit_r_environ()
```

Second, add your Census API key to your `.Renviron` file like so:

```
CENSUS_API_KEY="your_key_here"
```

This enables cpsR functions to automatically look up your key by running:

``` r
Sys.getenv("CENSUS_API_KEY")
```

## Example

```{r example, message=FALSE}
library(cpsR)
library(dplyr)
library(purrr)


# Simple use of the basic monthly CPS

sep21 <- get_basic(
  year = 2021,
  month = 9,
  vars = c("prpertyp", "prtage", "pemlr", "pwcmpwgt")
)

sep21

sep21 %>%
  filter(prpertyp == 2 & prtage >= 16) %>%
  summarize(
    pop16plus = sum(pwcmpwgt),
    employed = sum(pwcmpwgt[pemlr %in% 1:2])
  ) %>%
  mutate(epop_ratio = employed / pop16plus)


# Pulling multiple years of CPS ASEC microdata

asec <- map_dfr(2020:2021, get_asec, vars = c("h_year", "marsupwt"))

count(asec, h_year, wt = marsupwt)
```
