---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cpsR

<!-- badges: start -->

[![R-CMD-check](https://github.com/matt-saenz/cpsR/workflows/R-CMD-check/badge.svg)](https://github.com/matt-saenz/cpsR/actions)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)

<!-- badges: end -->

## Overview

Load [Current Population Survey (CPS)](https://www.census.gov/programs-surveys/cps/about.html) microdata from the Census Bureau Data API into R, including [basic monthly CPS](https://www.census.gov/data/datasets/time-series/demo/cps/cps-basic.html) and [CPS ASEC](https://www.census.gov/data/datasets/time-series/demo/cps/cps-asec.html) microdata.

Note: This product uses the Census Bureau Data API but is not endorsed or certified by the Census Bureau.

## Installation

To install cpsR, run the following code:

``` r
# install.packages("devtools")
devtools::install_github("matt-saenz/cpsR")
```

Next, register for a [Census API key](https://api.census.gov/data/key_signup.html) if you don't already have one. Once you have your key, store it in an environment variable named `CENSUS_API_KEY` for safe and easy use. You can do this in two steps:

1. Run `usethis::edit_r_environ()` to open your `.Renviron` file
2. Add `CENSUS_API_KEY="your_key_here"` to your `.Renviron` file

This allows cpsR functions to automatically add your key to Census API requests (via `Sys.getenv("CENSUS_API_KEY")`) and saves you from having to manually copy-paste your key around. Using env var `CENSUS_API_KEY` also allows you to avoid sharing your key in scripts you may share with others (or post on GitHub).

## Example

```{r example, message=FALSE}
library(cpsR)
library(dplyr)

basic <- get_basic(
  year = 2021,
  month = 3,
  vars = c("prpertyp", "prtage", "pemlr", "pwcmpwgt")
)

basic

basic %>%
  filter(prpertyp == 2 & prtage >= 16) %>%
  summarize(
    pop16plus = sum(pwcmpwgt),
    employed = sum(pwcmpwgt[pemlr %in% 1:2])
  ) %>%
  mutate(epop_ratio = employed / pop16plus)
```
